\section*{Acknowledgements}

% HOWTO: Please insert your acknowledgements of funding agencies etc. here. Please add a comment from which chapter they come. We will merge them to a nice text later.

\begin{itemize}

\item Nestor Armesto was supported by Ministerio de Ciencia e Innovaci\'on of Spain under projects FPA2014-58293-C2-1-P, FPA2017-83814-P and Unidad de Excelencia Mar\'{\i}a de Maetzu under project MDM-2016-0692, by Xunta de Galicia (Conseller\'{\i}a de Educaci\'on) within the Strategic Unit AGRUP2015/11, and by FEDER. %UPC/pPb
\item Raphaelle Bailhache acknowledges the support of the German Federal Ministry of Education and Research (BMBF).%Thermal radiation
\item Zvi Citron acknowledges the support by the ISRAEL SCIENCE FOUNDATION (grant No. 91/6491).%small A summary
\item Torsten Dahms acknowledges the support of the DFG cluster of excellence ``Origin and Structure of the Universe''.%Thermal radiation 
\item Stefan Floerchinger acknowledges the support by the DFG Collaborative Research Centre SFB 1225 (ISOQUANT).  %Light flavour
\item Iwona Grabowska-Bold acknowledges the support of the Polish National Science Centre grant DEC-2016/23/B/ST2/01409, of the AGH UST statutory tasks No. 11.11.220.01/4 within subsidy of the Ministry of Science and Higher Education, and of the PL-Grid Infrastructure.% Beyond / Gamma-Gamma
\item Ilkka Helenius was supported by the Carl Zeiss Foundation and the Academy of Finland, Project 308301.%UPC/pPb
\item Spencer Klein acknowledges the support of the U.S. DOE under contract number DE-AC02-05-CH11231.%Thermal radiation and UPC/pPb
\item Filip Krizek acknowledges the support of the Ministry of Education of the Czech Republic (grant InterExcellence LT17018) % Small systems
\item Mariola Klusek-Gawenda acknowledges the support of the Polish National Science Center Grant No. DEC-2014/15/B/ST2/02528.% Beyond / Gamma-Gamma
\item Nirbhay Kumar Behera acknowledges the National Research Foundation of Korea (NRF), Basic Science Research Program, funded by the Ministry of Education, Science and Technology (Grant No. NRF-2014R1A1A1008246). %Light flavour
\item Constantin Loizides is supported by the U.S. Department of Energy, Office of Science, Office of Nuclear Physics, under contract number DE-AC05-00OR22725.
\item Laure Massacrier acknowledges the support by CNRS under the grant RFBR/CNRS 18-52-15007.% Heavy flavour
\item Christoph Mayer acknowledges support from the Polish Ministry of Science and Higher Education and from the Polish National Science Center.%UPC/pPb
\item Petja Paakkinen has been supported by the Magnus Ehrnrooth Foundation.%UPC/pPb
\item Hannu Paukkunen is supported by his Academy-of-Finland project 308301.%UPC/pPb
\item Dmitri Peresunko acknowledges the support of the Russian Science Foundation grant 17-72-20234.%Thermal radiation
\item Ralf Rapp has been supported by the US National Science Foundation under grant number PHY-1614484, and in part by the ExtreMe Matter Institute EMMI at the GSI Helmholtzzentrum f\"{u}r Schwerionenforschung (Darmstadt,Germany).% Quarkonia
\item Kristof Redlich acknowledges the Polish National Science Center NCN under Maestro grant DEC-$\mathrm{2013/10/A/ST2/00106}$.  %Light flavour
\item Mark Strikman was supported by the U.S. Department of Energy, Office of Science, Office of Nuclear Physics, under Award No. DE-FG02-93ER40771.%UPC/pPb
\item Michael Weber acknowledges the support of the Austrian Academy of Sciences and the Nationalstiftung f\"ur Forschung, Technologie und Entwicklung, Austria.%Thermal radiation
\item Michael Winn acknowledges support from the European Research Council~(ERC) through the project EXPLORINGMATTER, funded by the ERC through a ERC-Consolidator-Grant.%UPC/pPb

\end{itemize}
