% don't remove the folling lines, and edit the defintion of \main if needed
\documentclass[../report.tex]{subfiles}
\providecommand{\main}{..}
\IfEq{\jobname}{\currfilebase}{\AtEndDocument{\biblio}}{}
\IfEq{\jobname}{\currfilebase}{\input{\main/commands}}{}
% until here


%<Notes>
% Dear contacts,

% Several chapters have started defining custom commands, and this lead to some compilation issues.

% I have now prepared a file commands.tex which is in the main folder. 
% This file is included when you compile the full report (report.tex), but also when you compile the separate sections.

% Please add your custom commands to this file. When you do that, please
% - check that a command does not already exist and can be used,
% - add a comment above your commands, e.g. "% heavy flavour"

% NB. pT is \pT (already predefined in a CERN report style file). Please replace in case you are using \pt

% Best regards,
% Jan Fiete
%</Notes>

% 05/09/18 09:31:59
% Hello John,

% Very good. In addition, we should write explictly in "your" chapter that we decided for the projections to use p = 1.5 ... 1.9 which converts into an increase of LNN of 8 ... 25.

% Best regards,
% Jan Fiete

\begin{document}
\section{Collider performance for heavy ions} 

\subsection{LHC performance with heavy ions up to Run 2}


This year's 4th one-month Pb-Pb run of the LHC will bring its Run~2 to an end and launch the hardware upgrades to the collider, and to the ALICE experiment, that should allow the full  ``HL-LHC'' heavy-ion performance to be delivered from 2021 onward.   
Indeed, if we compare the specifications of the 2004 LHC Design Report  \cite{lhc1}, for Pb-Pb collisions only with a peak luminosity of \lumival{1}{27},  much  of the upgraded performance is already in hand.   
Not only has that peak Pb-Pb luminosity goal been exceeded by a factor of 3.6, but the p-Pb collision mode---an upgrade beyond the initial design whose very feasibility was widely doubted---has yielded similarly high luminosity 
in multiple operating conditions  (see \cite{Jowett:2018yqk} and references therein). 
Table~\ref{tab:LHCparams} summarises the main parameters of the runs to date. 
Most recently, in 2017, the LHC has collided beams of Xe nuclei \cite{Schaumann:2018qat}, 
providing many of the new results presented in this conference.  
The goal for 2018 will be to complete the accumulation of goal 
of an integrated \PbPb\ 
luminosity of \qty{1}{nb^{-1}} to 
each of the ALICE, ATLAS and CMS experiments.  

	
\begin{table*}[t]
\caption{Representative simplified beam parameters 
at the start of the highest luminosity physics   fills, 
in conditions that lasted for $>\qty{5}{days}$,   
in each annual Pb-Pb and p--Pb run~\cite{ipac2011:Jowett:2011zz,Jowett:1492972,ipac2013:Jowett:1572994,ipac2016:PbPb2015,ipac2017:Jowett:2289686}.  
The original design values for Pb--Pb~\cite{lhc1} 
and p-Pb~\cite{Salgado:2011wc}  
and future upgrade Pb--Pb goals are also shown 
(in these columns the integrated luminosity goal is to be attained over the 4~P--Pb 
runs in the
{10-year periods} before and after 2020).
Peak and integrated luminosities are averages for ATLAS and CMS 
(ALICE being levelled).
The smaller luminosities delivered to LHCb from 2013--2016 and 
in the minimum-bias part of the run in 2016 are not shown.
Emittance and bunch length are RMS values. 
Single bunch parameters for \pPb\ or \Pbp\ runs 
are generally those of the Pb beam.
The series of runs with $\sqrt{\sNN}=\qty{5.02}{TeV}$ also included \pp\ reference runs, not shown here.
Design and record achieved nucleon-pair luminosities are \protect\fbox{boxed}, and some key \pPb\ parameters are 
set in \textcolor{red}{red type}, for easy comparison. 
The upgrade peak luminosity is reduced by a factor $\simeq3$ from its potential value by levelling.
\label{tab:LHCparams}}	
\centering
    {\small
        \input{accelerator/Run1Run2Table.tex}
    }
%%% $^{*}$ Pb and p beams reached 1577 and 4000~GeV/nucleon in 2012.
 
\end{table*} 


\subsection{Schedule for future LHC heavy-ion runs} 

Present nominal plan based on ALICE Letter of Intent 2012.

\subsection{\PbPb luminosity at HL-LHC}  \label{sec:HLPbPb} 

The LHC Design Report originally foresaw Pb-Pb collisions with a 
luminosity of  . 

A major step towards the future upgraded performance is the provision of sufficient total bunch intensity from the LHC injectors.  
A detailed specification of the requirements on the beams at LHC injection 
has been given 
\cite{HLLHCPbPbspec}
and is well on track to being fulfilled.  
The necessary single-bunch intensities have already been attained. 
Slip-stacking in SPS  
will be implemented in 2021. 


The High-Luminosity (HL-LHC) upgrade of the LHC will implement a number of 
hardware upgrades of the present LHC that will enable a substantial increase in p-p luminosity after 
Long Shutdown~3 (LS3).   

\subsection{Secondary beams from the IPs}

\subsubsection{Collimation and intensity limit}

\subsubsection{Upgrades during LHC Long Shutdown 2}
 
\subsection{\pPb\ at HL-LHC} 

 
\subsection{Lighter nuclei at HL-LHC} \label{sec:lightions}
 

The bunch intensity limits in the injectors depend largely on the ion charge which changes at the various stripping stages which must be optimised for space-charge limits, intra-beam scattering, efficiency of electron-cooling,   beam losses on residual gas and other effects in the ion source, Linac4,  LEIR, the PS and SPS.  
Given the uncertainties, the deliverable intensity for other species can only be determined after sufficient time spend commissioning and empirically optimising the many parameters and operating modes of the whole injector chain. 
To simplify present considerations, we postulate a simple form relating the number of ions per bunch, \Nb, to the well-established value for Pb beams  
\begin{equation}
{\Nb}(Z,A)   = {\Nb}(82,208){\left( {\frac{Z}{{82}}} \right)^{ - p}} 
\label{eq:pscaling}
\end{equation}
Fitting such an expression to the limited information~\cite{Manglunki:2016qzl} from the few species used for SPS fixed-target in recent years (since the commissioning of the present ECR ion source and LEIR) yields a value of the fit parameter \(p=1.9\).    
Beam quality requirements for fixed-target beams are, of course, less stringent than for injection into the collider.
Fitting to the first commissioning of Xe beams for the LHC~\cite{Schaumann:2018qat}, on the other hand, gives a much less optimistic \(p=0.75\).   
Although this was the only occasion where any other species than Pb was delivered to the collider, only the simplest version of the injection scheme was used and it is clear that, 
given time, significantly higher intensities could be achieved.  
We consider that  \(1.5\le p\le 1.9\) corresponds to a representative range of possibilities that could be 
realised in fully-prepared future operation.  


In addition, we make a number of simplifying assumptions to allow a simplified, yet meaningful, comparison between species 
\begin{itemize}
    \item  The     \emph{geometric} transverse beam emittances at the start of collisions will be equal to those of Pb beams~\cite{HLLHCPbPbspec}.
    This is justified, at least at the level of the LHC, since the scaling of intra-beam scattering with \(\Nb\), \(Z\) and \(A\), 
    given by  the parameter 
    $f_{\text{IBS}}\text{/(m Hz)}$  
    is generally smaller than for \Pb\ as long as 
    $p\lsim1.9$.  A similar scaling should hold in the injectors such as the SPS where intra-beam scattering may also blow up the emittances.  
    This ignores possible space-charge limits in the injectors which
    should also be considered once the appropriate stripping schemes and charge states have been defined. 
    
    \item Same filling scheme with \(k_c=\)
    \item No luminosity-levelling in any experiment.
     \item Fill length optimised for intensity evolution dominated by luminosity burn-off. 
     \item Equal operational efficiency of 50\%.     Following conventional practice for HL-LHC, the integrated luminosity for a 1-month run is estimated assuming back-to-back ideal fills of optimal length and a turn-around time of \qty{2.5}{h} between the end  of one fill and the resumption of ``Stable Beams''  for collisions in the next.
\end{itemize} 
The parameters are estimated using analytical approximations unlike the more elaborate simulations used in Section~\ref{sec:HLPbPb}.  
Together with the assumption that there is no luminosity levelling, these lead to a higher estimate of integrated luminosity in a one-month run.  
Nevertheless they can be used as a guide to the relative gain factors in integrated nucleon-nucleon luminosity by changing from Pb to a lighter nucleus. 

\begin{table}%% Table from HL-LHC species.nb for p=1.5 
\centering
{\small
	\input{accelerator/SpeciesTable1.tex}
	}
   \caption{Parameters and performance for a range of light nuclei with a moderately optimistic value of the scaling parameter \(p=1.5\) in \eqref{eq:pscaling}. }
\end{table}
%\vspace{1em}


\begin{table}%% Table from HL-LHC species.nb for p=1.5 
\centering
{\small
	\input{accelerator/SpeciesTable2.tex}
	}
   \caption{Parameters and performance for a range of light nuclei with an optimistic value of the scaling parameter \(p=1.9\) in \eqref{eq:pscaling}. }
\end{table} 
 
 Collimation \cite{Hermes:2016axvg}

\subsection{Short run for \pO}
\label{sec:pOrun} %% Hans Dembinski can refer to this section  

As discussed in Section~
\ref{sec:pOcosmic}, a  short \pO\ collision run is of interest for cosmic-ray physics.  
This could be scheduled with rapid set-up in LHC on the successful 
model of the 2012 \pPb\ run\cite{} 
which was later re-used in the 
2017 \XeXe\ run\cite{Schaumann:2018qat}.  
Each of those runs took  about 16~h of LHC operation time, 
including set-up and physics data-taking. 

Because oxygen is used as the carrier gas in the CERN heavy-ion source, 
the idea has been mooted that 
it may be possible to switch from \Pb\ to
\isotope{16}{O}{8+}  beams for the LHC, and back,  
somewhat more rapidly than other species.  
However it appears that the process is still not quick enough to allow, say, 
a short p--O run at the end of one of the annual Pb-Pb runs.  
A more practical possibility would be to schedule the run 
earlier in the year in order to provide time for 
the source to be switched back to Pb operation
afterwards for the one-month run later in the year.   
Commissioning of the O beam in the injectors for single-bunch injection into the LHC
would need to be scheduled in parallel with p-p operation in the weeks before the \pO\ run. 

To keep the LHC commissioning time on the order of 8~h, as in the 2017 Xe--Xe run,  
the total charge per beam should be limited to similar values.  
This would lead to a filling scheme with \qty{\simeq 10}{bunches}. 
A  probable disadvantage of this scheme is that the \pO\  run 
would have to be done with the currently operational \pp\ optics with larger \bstar, typically the injection value \bstarval{10}, at IP2.  
This would mean that the ALICE experiment would receive less integrated luminosity than ATLAS and CMS, 
as  occurred in the 2017 \XeXe\ run~]\cite{Schaumann:2018qat}.

\subsection{Heavy-ion performance of HE-LHC} 

For the moment assume same injected beams as HL-LHC 
in the absence of work on 
further possible upgrades to the injectors. 

Arguments from FCC Week in Amsterdam. 
2Modest factor in integrated luminosity.

BFPP power with Pb very high. 
Could be alleviated with lighter species.
Expect roughly similar species scaling to HL-LHC in Section~\ref{sec:lightions}.


\end{document}
