\subsubsection{Heavy-flavour angular correlations}
The investigation of $\PD \bar{\PD}$ correlations provides access to the initial correlation of ccbar quark production in the initial state in pPb and pp collisions. This is an important information to disentangle the contributions from different production channels. Furthermore, effects of decorrelation due to the interaction with the possibly created medium as well effects due to gluon saturation are accessible. We present a projection for the correlation function in the forward direction with LHCb which in particular interesting for saturation effects. The measurement will be statistically limited and only statistical uncertainties are shown in the plot. As demonstrated by the bin choice, a scan through the kinematics of the $\PD \APD$ pair is possible as a function of the D-meson $\pt$ providing differential information to test theoretical models with precision.  We provide the projection as well in the pp reference sample which can serve as baseline. These measurements can be complemented with correlation measurements of $\PDz \APDzero$  and $\PDz \PsDm$ as well $\PDz \PGLc^-$, $\PDp \PDm$, $\PDp\PsD-$ and $\PDp \PGLc^-$ as in publication~\cite{LHCb-PAPER-2012-003}.

\begin{itemize}
\item {Experiments:} B, D-jets, Di-Bjets
\item {Connection to theory:} constraints on mass dependence of in-medium splitting functions. Role of coherence (decoherence angle comparable to dead-cone angle).
\end{itemize}

\textbf{FIGURES}:
\begin{enumerate}
\item DDbar azimuthal correlation based on the pp measurement for pPb (LHCb)
\item D-tagged jets in PbPb, HFE-tagged jets (ALICE), B-jets (CMS)
\item D-jet correlations in PbPb (CMS)
\item Di-bjets (CMS)
\item Theory predictions together with projections
\end{enumerate}
